On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-16 manuscripts/PNAS/review/point_by_point_response.txt

TODO:

References for Cover Letter and Point-by-point Response:

(IHS10) Isely, Hillar, Sommer, Advances in neural information processing systems (NIPS), 2010.
(GS12) Ganguli, Sompolinksi, Annual review of neuroscience, 2012.
(TP14) Tillman, Pfetsch, IEEE Transactions on information processing, 2014.
(HS15) Hillar, Sommer, IEEE Transactions on information processing, 2015.
(M10) Morris I (2010), Advances in Mathematics 225(6):3425–3445.

@@

GENERAL COMMENTS

Overall, while I think that the manuscript makes a solid contribution, I believe is better suited to a more specialized journal. Unlike many journals, PNAS gives a criterion for significance: that the paper be within the top 10% of publications and should be of interest to a general scientific audience. Without compelling evidence to the contrary from the authors, it seems to me that the submission does not rise to that level.
-----------------

RESPONSE: We have included in our revised manuscript a more compelling exposition of the significance of our results and their relevance to the general scientific audience, and we thank the editor for the opportunity to address this concern shared by both reviewers.

-----------------
Since the result is an extension of the noise-free result of (HS15) to a noisy case, the authors should provide better justification for the significance of contributions. The main result seems to be based on an approach similar to (HS15), and there is little to no discussion of the computability and tightness of the sufficient condition for stability in the main results. Thus, the submission does not seem to be greatly more valuable than (HS15) and also not conclusive regarding the central problems of sparse linear coding and dictionary learning. That being said, the paper (with the modifications below) would certainly merit publication in very solid journal like IEEE Transactions on Information Theory, where it would be a median paper, in my opinion.
-----------------

@@

To be clear, the new perspective we take on the problem yields the following powerful conclusions beyond those of a straightforward extension of (HS15) to the noisy case:

1) An extension to the case where the number of dictionary elements is unknown: The results of (HS15) only apply to the case where the matrix B has the same number of columns as A. We forgo this assumption and show that B must have at least as many columns as A and contains (up to noise) a subset of the columns of A. The size of this subset depends on a simple relation between the number of columns of B and the regularity of the support-set hypergraph.

2) A significant reduction in sample complexity: To identify the n x m generating dictionary A, (HS15) require that data be sampled from every k-dimensional subspace spanned by the m columns of A (that is, {m \choose k} subspaces in total). We show that the data need only be sampled from m subspaces in total (e.g. those supported by consecutive intervals of length k in some cyclic order on [m]) and in some cases as few as 2\sqrt{m} (e.g. when m = k^2, take the supports to be the rows and columns of [m] arranged in a square grid). Moreover, if the size of the support set of reconstructing codes is known to be polynomial in m and k, then the pigeonholing argument in the proof of Thm. 1 requires only a polynomial number of samples distributed over a polynomial number of supports; thus, N is polynomial in m, k. This point was only hinted at in the Discussion section of our original submission, but we have included it in the updated manuscript to make clear the power of our approach over that taken in (HS15). We are aware of only one work treating combinatorial hypergraph designs in the setting of dictionary learning, "An incidence geometry approach to dictionary learning" (arxiv, non-published paper)

3) No spark condition requirement for the generating matrix A: One of the mathematically significant contributions made by (HS15) was to forgo the constraint that the recovery matrix B also satisfy the spark condition, in contrast to all previously known uniqueness results which (either explicitly or implicitly) made this assumption.  Our proof is powerful enough to show that, in fact, even the matrix A need not satisfy the spark condition to be identifiable! Rather, it need only be injective on the union of subspaces with supports that form sets in a hypergraph satisfying the singleton intersection property (SIP). (For example, consider the matrix A = [e_1, …, e_5, v] where v = e_1 + e_3 + e_5. and take H to be the set of all consecutive pairs of [m] arranged in cyclic order. Then A satisfies the assumptions of Thm. 1 without satisfying the spark condition.) We had omitted this point in our original submission of the manuscript to keep things simple, but have decided to include this fact in our resubmission. This also required us to redefine the restricted matrix lower bound L to be in terms of a hypergraph, which is an interesting object for further study in its own right.

@@

RESPONSE: Yes, thanks for this question. The condition on \ell is perhaps best understood as a condition on \bar m not being too large (given \ell) in order to guarantee the recovery of at least one dictionary element. We have updated the manuscript to reflect this, and the condition now reads “provided \bar m < \ell m / (\ell - 1)” where previously it read “provided p = \bar m − \ell( \bar m − m ) is positive”.

-----------------
-- The paper could also compare the sample complexity results on N with those given in the noise-free case. For example, Table I in reference (18) gives a comparison of sample complexity requirements for different conditions in the noise-free case. Where would the results of this paper stand in that table?

-- Otherwise, all the math seemed correct.
-----------------

RESPONSE: In brief, since all of the theorems and corollaries in (HS15) are corollaries of our theorems, all sample complexities improve by our work. The main point of difference between our statements about sample complexity and those of (HS15) is the assumed constraint on the underlying support set of sparse codes x_i. Our theory of combinatorial designs (the ‘singleton intersection property’) requires only that there be a sufficient number of x_i drawn from each support in a hypergraph satisfying the SIP, whereas the theory in (HS15) requires that a sufficient number of x_i be drawn from *every* support of size k in [m]. Here, we make an explicit comparison with Table I in (HS15) row by row, assuming for our results that sufficient data have been sampled from all supports in a hypergraph H satisfying the SIP:

ROW I. Our result here is |H|(k-1){m \choose k} + 1. We improve on the result in (HS15) by an exponential factor since our theory does not require that data be sampled with every possible support of size k in [m]; as noted in our manuscript, for every k < m there exists a hypergraph H with |H| = m that satisfies the SIP.

ROW 2. Our result here is again |H|(k-1){m \choose k} + 1 with *certainty*, not with *almost certainty* (i.e. probability 1) as in (HS15). Here the authors in (HS15) have applied probabilistic arguments to achieve an almost certain result with a sample complexity on the same order as that for which we have achieved a certain result by means of our theory of combinatorial designs.

ROW 3. We cannot make a direct comparison here because we have not calculated the probability that a random set of supports satisfy the SIP (see our response to a similar question by the second reviewer). We think this is an interesting problem for the community to solve; regardless, the result of (HS15) here is implied by our more general theory.

ROW 4. Our result here is |H|(k-1){m \choose k} + 1 with probability 1. Here is the only case where the sample complexity stated in (HS15) is technically better than ours, but it differs in flavor: their x_i are assumed to be distributed as (k+1) samples per support for every support of size k in [m], whereas ours supposes that (k-1){m \choose k} codes x_i are distributed over each support in some hypergraph H satisfying the SIP. Regardless, the (technically) lower result of (HS15) is still implied by our more general theory in the case \varepsilon = 0 with the addition of their probabilistic argument.

ROW 5: Again, we cannot make a direct comparison here for the reason stated in ROW 3.

While we appreciate this question, we feel that since our results improve in every case except for one noise-free technicality (ROW 4), and since the results of (HS15) are all entailed anyway by our more general theory, an explicit comparison such as that which we provided above would not be the best use of our limited space in the manuscript; although we are willing to add a discussion about this if necessary. Many thanks for your careful review of the mathematics. We have tried and continue to try to be as clear, concise, and correct as we possibly can to elevate this work into the top echelon of applied mathematical theory papers.

### 2017-08-16 manuscripts/PNAS/chazthm_pnas.tex
prop1

\begin{proposition}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse linear coding (e.g. \cite{rehnsommer2007, rozell2007neurally}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}

>>>>

\begin{proposition}\label{neuroprop}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse linear coding (e.g. \cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}
